---
title: "R Notebook"
output: html_notebook 
---

<!-- sourcing data and function file-->
```{r}
source_rmd <- function(x, ...) {
  library(knitr)
  options(knitr.duplicate.label = 'allow')
  source(purl(x, output = tempfile()), ...)
}
source_rmd("../code/s1_Load_Preprocess_MSKCC.Rmd")
```

<!-- Run s0 and s1 to load and preprocess the data and functions required-->
<!-- Function required for figure1-->
<!--Analysis reviewer 3 - R4 and R5-->
<!--Update: use of argument "th" in function for the making the percentile based TMB threshold instead of 10 mut/MB -->
```{r}
Calculate_TMB_Response_association_with_CI<-function(infunc_df= mskcc_combined,
                                             COI="HNSC",
                                             response_measure='OS'
                                             # ,th= ratio_TMB_H_vs_L *100
                                             ){
  infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type_acronym == COI)),]
  #infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type == COI)),]
  
  if(response_measure=='ORR'){
    infunc_df_tmb=infunc_df_tmb[!is.na(infunc_df_tmb$ORR),]
    # threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
    # Contigency matrix
    cont_matrix=table( infunc_df_tmb$TMB >= 10, infunc_df_tmb$ORR) ## use threshold instead of 10 
    # Perform fisher test
    fisher_test_results_raw=fisher.test(cont_matrix)
    to_return=c(effect_size=unlist(fisher_test_results_raw$estimate),
                P=fisher_test_results_raw$p.value,
                lower_ci=fisher_test_results_raw$conf.int[1],
                upper_ci=fisher_test_results_raw$conf.int[2])
  } else if(response_measure =='OS') {
    infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$Overall_survival_months),]
    # threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
    
    # Perform coxph test
    coxph_model <- coxph(Surv(time = Overall_survival_months,
                              event = Overall_survival_status) ~ (TMB >= 10) , ## use threshold instead of 10 
                         data = infunc_df_tmb)
    to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
                P=summary(coxph_model)$coefficients[1,c(5)],
                lower_ci= summary(coxph_model)$conf.int[1,c(3)],
                upper_ci= summary(coxph_model)$conf.int[1,c(4)])
  }  else if(response_measure=='PFS') {
    infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$PFS_months),]
    # threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
    
    coxph_model <- coxph(Surv(time = PFS_months,
                              event = PFS_status) ~ (TMB>=10) ,  ## use threshold instead of 10 
                         data = infunc_df_tmb)
    to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
                P=summary(coxph_model)$coefficients[1,c(5)],
                lower_ci= summary(coxph_model)$conf.int[1,c(3)],
                upper_ci= summary(coxph_model)$conf.int[1,c(4)])
  }
  names(to_return)=c('effect_size', 'P', 'lower_ci', 'upper_ci')
  to_return
}
```

```{r}

allcancer_types=levels(mskcc_combined$Cancer_Type_acronym)
OS_cox_results= do.call(rbind, lapply(allcancer_types, function(x)
  err_handle(Calculate_TMB_Response_association_with_CI(infunc_df= mskcc_combined,
                                                COI=x,
                                                response_measure='OS'
                                                # ,th = ratio_TMB_H_vs_L* 100
                                                ))))
OS_cox_results<- data.frame(OS_cox_results)
rownames(OS_cox_results)=allcancer_types

## PFS
allcancer_types=levels(mskcc_combined$Cancer_Type_acronym)
PFS_cox_results= do.call(rbind, lapply(allcancer_types, function(x)
  err_handle(Calculate_TMB_Response_association_with_CI(infunc_df= mskcc_combined,
                                                COI=x,
                                                response_measure='PFS'
                                                # ,th = ratio_TMB_H_vs_L* 100
                                                ))))
PFS_cox_results<- data.frame(PFS_cox_results)
rownames(PFS_cox_results)=allcancer_types

## ORR
ORR_cox_results= do.call(rbind, lapply(allcancer_types, function(x)
  err_handle(Calculate_TMB_Response_association_with_CI(infunc_df= mskcc_combined,
                                                COI=x,
                                                response_measure='ORR'
                                                # , th = ratio_TMB_H_vs_L* 100
                                                ))))
ORR_cox_results<- data.frame(ORR_cox_results)
rownames(ORR_cox_results)=allcancer_types
```

<!--Analysis reviewer 2 : adding the sample size which have high TMB-->
```{r}
TMB_level_by_cancer_type = table(mskcc_combined$Cancer_Type_acronym,
                                 (mskcc_combined$TMB >10))
colnames(TMB_level_by_cancer_type) = c("less_than_10", "greater_than_10")
ratio_T_F= data.frame(greater_than_10= TMB_level_by_cancer_type[,"greater_than_10"], 
                      less_than_10= TMB_level_by_cancer_type[,"less_than_10"],
                      ratio= TMB_level_by_cancer_type[,"greater_than_10"] / 
                        (TMB_level_by_cancer_type[,"greater_than_10"]+TMB_level_by_cancer_type[,"less_than_10"]))

mskcc_combined_orr= mskcc_combined[!is.na(mskcc_combined$ORR),]
TMB_level_by_cancer_type_ORR = table(mskcc_combined_orr$Cancer_Type_acronym,
                                     (mskcc_combined_orr$TMB >10))

colnames(TMB_level_by_cancer_type_ORR) = c("less_than_10", "greater_than_10")
ratio_T_F_OR_PFS= data.frame(greater_than_10= TMB_level_by_cancer_type_ORR[,"greater_than_10"], 
                      less_than_10= TMB_level_by_cancer_type_ORR[,"less_than_10"],
                      ratio= TMB_level_by_cancer_type_ORR[,"greater_than_10"] / 
                        (TMB_level_by_cancer_type_ORR[,"greater_than_10"]+TMB_level_by_cancer_type_ORR[,"less_than_10"]))


# proportion_TMB = ggplot(ratio_T_F, aes(x=reorder(rownames(ratio_T_F), ratio),
#                                            y= ratio, fill= ratio))+
#   geom_bar(stat = "identity")+
#   theme_bw(base_size = 15)+
#   labs(x='Cancer types',
#        y='Proportion of TMB(>10)')+
#   coord_flip()

# ggsave('/Users/neelamf2/Documents/GitHub/Immune_determinant_for_power_of_TMB/results/Proportion_TMB_greater_10.pdf', proportion_TMB, height=5, width = 6, units = 'in', dpi = 500)
```

<!--Analysis reviewer 2 : R2, remove PAAD-->
<!--#################################################################
# Figure 1A: Forest plot for OS vs TMB >= 10 all cancer type
################################################################# --->
```{r}
all_cancertype = samples_by_cancer_type[,1][samples_by_cancer_type[,2]> 1]
Figure_1A_plot_df= OS_cox_results
Figure_1A_plot_df= Figure_1A_plot_df[all_cancertype,]
Figure_1A_plot_df= na.omit(Figure_1A_plot_df)
Figure_1A_plot_df=Figure_1A_plot_df[ order(Figure_1A_plot_df$effect_size, decreasing = F),]

### in case of percentile : 10,
Figure_1A_plot_df= Figure_1A_plot_df[!(rownames(Figure_1A_plot_df)%in% c('KIRC'
                                                                         #,'PAAD'
                                                                         )),]
colnames(Figure_1A_plot_df)[2] = "pval"
### number of Samples 
no_of_samples_os= samples_by_cancer_type[samples_by_cancer_type[,2]> 1,]
no_of_samples_os = no_of_samples_os[match(rownames(Figure_1A_plot_df),no_of_samples_os$Cancer_Type_acronym),]

ratio_T_F= ratio_T_F[intersect(rownames(Figure_1A_plot_df),rownames(ratio_T_F)),]
##############
#fdr correction
Figure_1A_plot_df$fdr_corrected_pval = p.adjust(Figure_1A_plot_df$pval/2, method = "fdr")


### Adding columns in the forest plot ###

tabletext_OS=cbind(c("Cancer Type",
                      rownames(Figure_1A_plot_df)),
                    c("Samples", 
                      no_of_samples_os$ID),
                    c("# TMB>=10 mut/MB",
                      round(as.numeric(ratio_T_F$greater_than_10),2)),
                    c("pval",
                      format(Figure_1A_plot_df$pval/2, scientific = T, digits = 2)),
                   c("fdr_pval",
                     format(Figure_1A_plot_df$fdr_corrected_pval, scientific = T, digits = 2)),
                   c("lower CI",
                   round(log(Figure_1A_plot_df$lower_ci,10), 2)),
                   c("upper CI",
                   round(log(Figure_1A_plot_df$upper_ci,10), 2))
                   )

## data
HR_OS=(Figure_1A_plot_df)
### styles for color according to pval
styles_OS <- fpShapesGp(
  lines = list(gpar(col ="blue"),
               gpar(col = ifelse((HR_OS$pval/2)[1] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[2] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[3] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[4] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[5] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[6] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[7] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[8] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[9] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[10] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[11] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[12] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_OS$pval/2)[13] <0.05, "blue", "red"))
               # ,gpar(col = ifelse((HR_OS$pval/2)[14] <0.05, "blue", "red"))
  ),
  box = list(gpar(fill = "blue"),
             gpar(fill = ifelse((HR_OS$pval/2)[1] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[2] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[3] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[4] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[5] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[6] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[7] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[8] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[9] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[10] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[11] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[12] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_OS$pval/2)[13] <0.05, "blue", "red"))
             # ,gpar(fill = ifelse((HR_OS$pval/2)[14] <0.05, "blue", "red"))
             )
)
pdf('../results/panel_1_A.pdf', width = 15)
  
forestplot(tabletext_OS,
           boxsize = .30, # We set the box size to better visualize the type
           line.margin = .30, # We need to add this to avoid crowding
           mean = c(NA,
                    log(HR_OS$effect_size,10)),
           lower =c(NA,
                   log(HR_OS$lower_ci, 10)),
           upper = c(NA,
                    log(HR_OS$upper_ci,10)),
           clip=c(-1,1),
          lty.ci = c(1),
           ci.vertices = TRUE,
           graph.pos=4, # position of the box plot
           graphwidth = unit(12,"cm"),
           colgap = unit(1.2,"mm"),
           col=fpColors(box= styles_OS$box[1]),
           xlab="Log(10) HR",
           ylab= "Cancer Types",
           shapes_gp= styles_OS,
           txt_gp = fpTxtGp(label = list(gpar(fontface = 1,cex= 1.5),
                                         gpar(fontface = 1, cex= 1.5), 
                                         gpar(fontface = 1,cex=1.5)),
                            ticks=gpar(cex=1.25),
                            xlab = gpar(cex=1.25),
                            title = gpar(fontface = 2,cex= 1.48)),
          grid = structure(c(-1), gp = gpar(lty = 1, col = "black")),
           title = 'Power of TMB of OS between TMB high(>= 20 percentile) and TMB low ')

dev.off()
```

<!--#################################################################
# Figure 1B: Forest plot for ORR vs TMB cancer type with sample more than 50/all
################################################################# --->
```{r}
cancertype = samples_by_cancer_type[,1][samples_by_cancer_type[,2]> 1]
ORR_cox_results
Figure_1B_plot_df= ORR_cox_results[rownames(ORR_cox_results) %in% cancertype,]
Figure_1B_plot_df= Figure_1B_plot_df[!is.na(Figure_1B_plot_df$effect_size) & 
                                     is.finite(Figure_1B_plot_df$effect_size),]

Figure_1B_plot_df= Figure_1B_plot_df[!(rownames(Figure_1B_plot_df) %in% c("STAD",
                                                                          'PAAD',
                                                                          'GBM', ## in case of percentile
                                                                          "Unknown-Primary")),]
Figure_1B_plot_df=Figure_1B_plot_df[order(Figure_1B_plot_df$effect_size,
                                                                  decreasing = F),]

no_of_samples= samples_each_cancer_type[samples_by_cancer_type$ID >1,]
no_of_samples= no_of_samples[match(rownames(Figure_1B_plot_df),no_of_samples$Cancer_Type_acronym),]

ratio_T_F_orr= ratio_T_F_OR_PFS[intersect(rownames(Figure_1B_plot_df),rownames(ratio_T_F_OR_PFS)),]
colnames(Figure_1B_plot_df)[2]= "pval"

##############
#fdr correction
Figure_1B_plot_df$fdr_corrected_pval = p.adjust(Figure_1B_plot_df$pval/2, method = "fdr")

### plot ###
tabletext_ORR= cbind(c("Cancer Type",
                   rownames(Figure_1B_plot_df)), 
                 c("# patient",
                   no_of_samples$ID),
                 c("# TMB>=10 mut/MB",
                      round(as.numeric(ratio_T_F_orr$greater_than_10),2)),
                 c("pval",
                   format(Figure_1B_plot_df$pval/2, scientific = T, digits = 2)),
                 c("fdr",
                   format(Figure_1B_plot_df$fdr_corrected_pval, scientific = T, digits = 2)),
                 c("lower",
                   round(log(Figure_1B_plot_df$lower_ci,10), 2)),
                 c("upper",
                   round(log(Figure_1B_plot_df$upper_ci,10), 2))
                 )

HR_ORR=Figure_1B_plot_df
styles <- fpShapesGp(
  lines = list(gpar(col = ifelse((HR_ORR$pval/2) <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[1] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[2] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[3] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[4] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[5] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[6] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[7] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[8] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[9] <0.05, "blue", "red")),
                                gpar(col = ifelse((HR_ORR$pval/2)[10] <0.05, "blue", "red"))
               # ,
               #                  gpar(col = ifelse((HR_ORR$pval/2)[11] <0.05, "blue", "red")),
               #                  gpar(col = ifelse((HR_ORR$pval/2)[12] <0.05, "blue", "red")),
               # gpar(col = ifelse((HR_ORR$pval/2)[13] <0.05, "blue", "red"))
               ),
  box = list(gpar(fill = ifelse((HR_ORR$pval/2) <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[1] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[2] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[3] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[4] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[5] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[6] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[7] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[8] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[9] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_ORR$pval/2)[10] <0.05, "blue", "red"))
             # ,gpar(fill = ifelse((HR_ORR$pval/2)[11] <0.05, "blue", "red"))
             # ,gpar(fill = ifelse((HR_ORR$pval/2)[12] <0.05, "blue", "red")),
             # gpar(fill = ifelse((HR_ORR$pval/2)[13] <0.05, "blue", "red"))
             )
)
pdf('../results/panel_1_B.pdf',
     width = 15)

forestplot(tabletext_ORR,
           boxsize = .30, # We set the box size to better visualize the type
           line.margin = .30, # We need to add this to avoid crowding
           mean = c(NA,
                    log(HR_ORR$effect_size,10)),
           lower =c(NA,
                    log(HR_ORR$lower_ci,10)),
           upper = c(NA,
                     log(HR_ORR$upper_ci,10)),
           clip=c(-1,5),
           lty.ci = c(1),
           ci.vertices = TRUE,
           graph.pos=4, # position of the box plot
           graphwidth = unit(12,"cm"),
           colgap = unit(1.2,"mm"),
           col=fpColors(box= styles$box[1]),
           xlab="Log (10) Odds Ratio",
           ylab= "Cancer Types",
           shapes_gp = styles,
           txt_gp = fpTxtGp(label = list(gpar(fontface = 1,cex= 1.5),
                                         gpar(fontface = 1, cex= 1.5), 
                                         gpar(fontface = 1,cex=1.5)),
                            ticks=gpar(cex=1.25),
                            xlab = gpar(cex=1.25),
                            title = gpar(fontface = 2,cex= 1.48)),
           grid = structure(c(-1), gp = gpar(lty = 1, col = "black")),
           title = 'Power of TMB of Recists response between high and low TMB group')

dev.off()
```

<!-- #################################################################
# Extended Figure 1: Forest plot for PFS vs TMB cancer type with sample more than 50/ all
#################################################################-->
```{r}
cancertype = samples_each_cancer_type[,1][samples_each_cancer_type[,2]> 1]

Ex_figure_1_plot_df= PFS_cox_results
Ex_figure_1_plot_df= Ex_figure_1_plot_df[rownames(Ex_figure_1_plot_df) 
                                                   %in% cancertype,]# remove cancer type less than 50
Ex_figure_1_plot_df= na.omit(Ex_figure_1_plot_df)
Ex_figure_1_plot_df= Ex_figure_1_plot_df[!(rownames(Ex_figure_1_plot_df) %in% c("STAD",
                                                                                # 'KIRC',
                                                                                'SCLC',
                                                                                "Unknown-Primary")),]
Ex_figure_1_plot_df=Ex_figure_1_plot_df[
  order(Ex_figure_1_plot_df$effect_size, decreasing = F),]

### number of Samples 
no_of_samples_PFS= samples_each_cancer_type[samples_by_cancer_type$ID >1,]
no_of_samples_PFS = no_of_samples_PFS[match(rownames(Ex_figure_1_plot_df),no_of_samples_PFS$Cancer_Type_acronym),]

ratio_T_F_pfs= ratio_T_F_OR_PFS[intersect(rownames(Ex_figure_1_plot_df),rownames(ratio_T_F_OR_PFS)),]

colnames(Ex_figure_1_plot_df)[2] = "pval"
Ex_figure_1_plot_df$fdr_corrected_pval = p.adjust(Ex_figure_1_plot_df$pval/2, method = "fdr")

### plot ###
tabletext_PFS= cbind(c("Cancer Type",
                      rownames(Ex_figure_1_plot_df)),
                    c("# Samples", 
                      no_of_samples_PFS$ID),
                    c("# TMB>=10 mut/MB",
                      round(as.numeric(ratio_T_F_pfs$greater_than_10),2)),
                    c("pval",
                      format(Ex_figure_1_plot_df$pval/2, scientific = T, digits = 2)),
                    c("fdr",
                      format(Ex_figure_1_plot_df$fdr_corrected_pval, scientific = T, digits = 2)),
                    c("lower CI",
                      round(log(Ex_figure_1_plot_df$lower_ci,10), 2)),
                    c("upper CI",
                      round(log(Ex_figure_1_plot_df$upper_ci,10), 2))
                    )

## data
HR_PFS=Ex_figure_1_plot_df

######################
#styles for color according to pval
#####################
styles_PFS <- fpShapesGp(
  lines = list(gpar(col ="Blue"),
               gpar(col = ifelse((HR_PFS$pval/2)[1] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[2] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[3] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[4] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[5] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[6] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[7] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[8] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[9] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[10] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[11] <0.05, "blue", "red")),
               gpar(col = ifelse((HR_PFS$pval/2)[12] <0.05, "blue", "red"))
               , gpar(col = ifelse((HR_PFS$pval/2)[13] <0.05, "blue", "red"))
               # , gpar(col = ifelse((HR_PFS$pval/2)[14] <0.05, "blue", "red"))
  ),
  box = list(gpar(fill = "blue"),
             gpar(fill = ifelse((HR_PFS$pval/2)[1] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[2] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[3] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[4] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[5] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[6] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[7] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[8] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[9] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[10] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[11] <0.05, "blue", "red")),
             gpar(fill = ifelse((HR_PFS$pval/2)[12] <0.05, "blue", "red"))
             ,gpar(fill = ifelse((HR_PFS$pval/2)[13] <0.05, "blue", "red"))
             # ,gpar(fill = ifelse((HR_PFS$pval/2)[14] <0.05, "blue", "red"))
  ))

#######################
#Forest plot
######################
pdf('../results/Extended_panel_1.pdf', width = 15)
forestplot(tabletext_PFS,
           boxsize = .30, # We set the box size to better visualize the type
           line.margin = .30, # We need to add this to avoid crowding
           mean = c(NA,
                    log(HR_PFS$effect_size,10)),
           lower =c(NA,
                    log(HR_PFS$lower_ci,10)),
           upper = c(NA,
                     log(HR_PFS$upper_ci,10)),
           clip=c(-1,1),
           lty.ci = c(1),
           ci.vertices = TRUE,
           graph.pos=4, # position of the box plot
           graphwidth = unit(12,"cm"),
           colgap = unit(1.2,"mm"),
           col=fpColors(box= styles_PFS$box[1]),
           xlab="Log (10) Odd Ratio",
           ylab= "Cancer Types",
           shapes_gp = styles_PFS,
           txt_gp = fpTxtGp(label = list(gpar(fontface = 1,cex= 1.5),
                                         gpar(fontface = 1, cex= 1.5), 
                                         gpar(fontface = 1,cex=1.5)),
                            ticks=gpar(cex=1.25),
                            xlab = gpar(cex=1.25),
                            title = gpar(fontface = 2,cex= 1.48)),
           grid = structure(c(-1), gp = gpar(lty = 1, col = "black")),
           title = 'TMB Power of Progression Free response between high and low TMB group')
dev.off()
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
