---
title: "R Notebook"
## title: "Data preprocessing and function"
output: html_notebook
---
<!-- Libraries -->
```{r}
# install.packages("https://cran.r-project.org/src/contrib/Archive/rlang/rlang_0.4.10.tar.gz", repos = NULL, type="source")
# remove.packages('xfun')
# install.packages('xfun')
# packageVersion('xfun')
require(ggrepel)
# library(forestplot)
library(corrplot)
require("survival")
require(statar)
library(ggplot2)
library(ggpubr)
library(readxl)
library(tidyr)
require("survival")
require("survminer")
library(caret)
# setwd('../data/')
```

<!--Load MSKCC combined data -->
```{r}
<<<<<<< HEAD
mskcc_combined= read.csv("../data/mskcc_combined_updated.csv")

# ########################################
### Analysis: reviewer 2: R3 - Repeat Analysis excluding PAAD #################
# mskcc_combined= mskcc_combined[!mskcc_combined$Cancer_Type_acronym == "PAAD",]
# ########################################

mskcc_combined$Cancer_Type_acronym=factor(mskcc_combined$Cancer_Type_acronym)
samples_by_cancer_type = aggregate(ID ~ Cancer_Type_acronym,mskcc_combined, length)
=======
mskcc_combined= read.csv("../data/mskcc_combined.csv")

# ########################################
### Analysis: reviewer 2: R3 - Repeat Analysis excluding PAAD #################
# mskcc_combined= mskcc_combined[!mskcc_combined$Cancer_Type_acronym == "PAAD",]
# ########################################

mskcc_combined$Cancer_Type_acronym=factor(mskcc_combined$Cancer_Type_acronym)
table(mskcc_combined$Cancer_Type_acronym)
samples_by_cancer_type = aggregate(ID ~ Cancer_Type_acronym,mskcc_combined, length) 
>>>>>>> c35299f9b79869f425eb086bc1858c00695e4909
cancerTypes_of_Interest = samples_by_cancer_type[,1][samples_by_cancer_type[,2]> 1]

## removing cancer types where PFS/ ORR is NA for figure 1B and ex figure 1
mskcc_df_ORR_PFS= mskcc_combined[!(is.na(mskcc_combined$PFS_status)),]
<<<<<<< HEAD
samples_each_cancer_type = aggregate(ID ~ Cancer_Type_acronym,mskcc_df_ORR_PFS, length)
cancerTypes_each_Interest = samples_each_cancer_type[,1][samples_each_cancer_type[,2]> 1]
=======
samples_each_cancer_type = aggregate(ID ~ Cancer_Type_acronym,mskcc_df_ORR_PFS, length) 
cancerTypes_of_Interest = samples_by_cancer_type[,1][samples_by_cancer_type[,2]> 1]
>>>>>>> c35299f9b79869f425eb086bc1858c00695e4909
```
<!-- funtion Error handle -->
```{r}
err_handle<-function(x)
{ tryCatch(x, error=function(e){NA}) 
  }
```
<!--Analysis: reviewer 3- R5, finding equal number of sample when TMB >= 10 mut/MB, using percentile-->
```{r}
ratio_TMB_H_vs_L= nrow(mskcc_combined[mskcc_combined$TMB > 10,])/length(mskcc_combined$TMB)

TMB_level_by_cancer_type=table(mskcc_combined$Cancer_Type,(mskcc_combined$TMB >= 10))
colnames(TMB_level_by_cancer_type) = c("less_than_10", "greater_than_10")
ratio_T_F= data.frame(greater_than_10= TMB_level_by_cancer_type[,"greater_than_10"], 
                      less_than_10= TMB_level_by_cancer_type[,"less_than_10"],
                      ratio= TMB_level_by_cancer_type[,"greater_than_10"] / 
                        (TMB_level_by_cancer_type[,"greater_than_10"]+TMB_level_by_cancer_type[,"less_than_10"])
                      )
sum(ratio_T_F$greater_than_10)


### Sanity check####
check_sample_equal_in_thresholds<-function(infunc_df= mskcc_combined,
                                           COI="NSCLC",
                                           th= ratio_TMB_H_vs_L){
  infunc_df_tmb= infunc_df[infunc_df$Cancer_Type_acronym == COI,]
  
  threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(.266))]
  TMB_level_by_cancer_type = table(infunc_df_tmb$Cancer_Type, infunc_df_tmb$TMB >= threshold)
  c(less_than_threshold= TMB_level_by_cancer_type[,1],
    greater_than_threshold= TMB_level_by_cancer_type[,2], 
    threshold=threshold)
}

allcancer_types=levels(mskcc_combined$Cancer_Type_acronym)
samples_equal_in_threshold_each_cancer_type = do.call(rbind, lapply(allcancer_types, function(x)
  err_handle(check_sample_equal_in_thresholds(infunc_df= mskcc_combined,
                                                COI=x,
                                                th= ratio_TMB_H_vs_L)) ))

rownames(samples_equal_in_threshold_each_cancer_type)=allcancer_types
sum(samples_equal_in_threshold_each_cancer_type[,'greater_than_threshold'])
```
```{r}
table(mskcc_combined$stage)

```

<!--Revised Figure 2 (Manuscript)-->
<!--Analysis: reviewer 2, comment 2- Checking the sex association with tmb power-->
<!--Analysis: reviewer 2- comment 9(ii) Removing the ORR from the Immune factor dataframe-->
```{r}
M_F_Ratio= table(mskcc_combined$Cancer_Type_acronym, mskcc_combined$Sex)
M_F_Ratio_df= data.frame(male= M_F_Ratio[,"Male"],
                         female= M_F_Ratio[,"Female"],
                         ratio= M_F_Ratio[,"Male"]/M_F_Ratio[,"Female"])


####################################
#Add Sex ratio (male to female) in joo's immune factor dataframe for the Analysis R7-->
####################################
immune_factor[na.omit(match(rownames(M_F_Ratio_df),rownames(immune_factor))),"sex"] = M_F_Ratio_df$ratio
immune_factor = immune_factor[,!(colnames(immune_factor) %in% c("ORR"))]
```
<!--Analysis: reviewer 3- R5, finding equal number of sample when TMB >= 10 mut/MB, using percentile-->
```{r}


ratio_TMB_H_vs_L= nrow(mskcc_combined[mskcc_combined$TMB > 10,])/length(mskcc_combined$TMB)

TMB_level_by_cancer_type=table(mskcc_combined$Cancer_Type,(mskcc_combined$TMB >= 10))
colnames(TMB_level_by_cancer_type) = c("less_than_10", "greater_than_10")
ratio_T_F= data.frame(greater_than_10= TMB_level_by_cancer_type[,"greater_than_10"], 
                      less_than_10= TMB_level_by_cancer_type[,"less_than_10"],
                      ratio= TMB_level_by_cancer_type[,"greater_than_10"] / 
                        (TMB_level_by_cancer_type[,"greater_than_10"]+TMB_level_by_cancer_type[,"less_than_10"])
                      )
sum(ratio_T_F$greater_than_10)



### Sanity check####
check_sample_equal_in_thresholds<-function(infunc_df= mskcc_combined,
                                           COI="NSCLC",
                                           th= ratio_TMB_H_vs_L){
  infunc_df_tmb= infunc_df[infunc_df$Cancer_Type_acronym == COI,]
  
  threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(.266))]
  TMB_level_by_cancer_type = table(infunc_df_tmb$Cancer_Type, infunc_df_tmb$TMB >= threshold)
  c(less_than_threshold= TMB_level_by_cancer_type[,1],
    greater_than_threshold= TMB_level_by_cancer_type[,2], 
    threshold=threshold)
}

allcancer_types=levels(mskcc_combined$Cancer_Type_acronym)
samples_equal_in_threshold_each_cancer_type = do.call(rbind, lapply(allcancer_types, function(x)
  err_handle(check_sample_equal_in_thresholds(infunc_df= mskcc_combined,
                                                COI=x,
                                                th= ratio_TMB_H_vs_L)) ))

rownames(samples_equal_in_threshold_each_cancer_type)=allcancer_types
sum(samples_equal_in_threshold_each_cancer_type[,'greater_than_threshold'])
```

<!-- Functions needed-->
<!--Update: in original analysis we did not use the argument th, please uncomment the former function (code) and comment the latter function (code) when you perform the analysis based on TMB high >= 10 mut/MB -->

<!--Analysis: Reviewer 3 - R4, repeat the analysis at different percentile (10, 20 percentile)-->
<!--Update: use of argument "th" in function for the making the percentile based TMB threshold instead of 10 mut/MB -->

```{r}
<<<<<<< HEAD
############################################
#uncomment this function for the original analysis
##### Calculates TMB and Response association
############################################
=======
##### Calculates TMB and Response assocaition

# Calculate_TMB_Response_association<-function(infunc_df= mskcc_combined,
#                                              COI="NSCLC",
#                                              response_measure='ORR'){
#   infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type_acronym == COI)),]
#   #infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type == COI)),]
#   
#   if(response_measure=='ORR'){
#     infunc_df_tmb=infunc_df_tmb[!is.na(infunc_df_tmb$ORR),]
#     
#     # Contigency matrix
#     cont_matrix=table(infunc_df_tmb$TMB >= 10, infunc_df_tmb$ORR)
#     # Perform fisher test
#     fisher_test_results_raw=fisher.test(cont_matrix)
#     to_return=c(effect_size=unlist(fisher_test_results_raw$estimate),
#                 P=fisher_test_results_raw$p.value)
#   } else if(response_measure =='OS') {
#     infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$Overall_survival_months),]
#     
#     # Perform coxph test
#     coxph_model <- coxph(Surv(time = Overall_survival_months,
#                               event = Overall_survival_status) ~ (TMB >= 10) ,
#                          data = infunc_df_tmb)
#     to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
#                 P=summary(coxph_model)$coefficients[1,c(5)])
#   }  else if(response_measure=='PFS') {
#     infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$PFS_months),]
#     
#     coxph_model <- coxph(Surv(time = PFS_months,
#                               event = PFS_status) ~ (TMB >= 10) ,
#                          data = infunc_df_tmb)
#     to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
#                 P=summary(coxph_model)$coefficients[1,c(5)])
#   }
#   names(to_return)=c('effect_size', 'P')
#   to_return
# }
>>>>>>> c35299f9b79869f425eb086bc1858c00695e4909

Calculate_TMB_Response_association<-function(infunc_df= mskcc_combined,
                                             COI="NSCLC",
                                             response_measure='ORR',
                                             th= ratio_TMB_H_vs_L*100){
  infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type_acronym == COI)),]
  #infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type == COI)),]

  if(response_measure=='ORR'){
    infunc_df_tmb=infunc_df_tmb[!is.na(infunc_df_tmb$ORR),]
<<<<<<< HEAD

    # Contigency matrix
    cont_matrix=table(infunc_df_tmb$TMB >= 10, infunc_df_tmb$ORR)
=======
    threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
    
    # Contigency matrix
    cont_matrix=table(infunc_df_tmb$TMB >= threshold, infunc_df_tmb$ORR)
>>>>>>> c35299f9b79869f425eb086bc1858c00695e4909
    # Perform fisher test
    fisher_test_results_raw=fisher.test(cont_matrix)
    to_return=c(effect_size=unlist(fisher_test_results_raw$estimate),
                P=fisher_test_results_raw$p.value)
  } else if(response_measure =='OS') {
    infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$Overall_survival_months),]
<<<<<<< HEAD

=======
    threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
    
>>>>>>> c35299f9b79869f425eb086bc1858c00695e4909
    # Perform coxph test
    coxph_model <- coxph(Surv(time = Overall_survival_months,
                              event = Overall_survival_status) ~ (TMB >= threshold) ,
                         data = infunc_df_tmb)
    to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
                P=summary(coxph_model)$coefficients[1,c(5)])
  }  else if(response_measure=='PFS') {
    infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$PFS_months),]
<<<<<<< HEAD

    coxph_model <- coxph(Surv(time = PFS_months,
                              event = PFS_status) ~ (TMB >= 10) ,
=======
    threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
    
    coxph_model <- coxph(Surv(time = PFS_months,
                              event = PFS_status) ~ (TMB >= threshold) ,
>>>>>>> c35299f9b79869f425eb086bc1858c00695e4909
                         data = infunc_df_tmb)
    to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
                P=summary(coxph_model)$coefficients[1,c(5)])
  }
  names(to_return)=c('effect_size', 'P')
  to_return
}

# Calculate_TMB_Response_association<-function(infunc_df= mskcc_combined,
#                                              COI="NSCLC",
#                                              response_measure='ORR',
#                                              th= 20#ratio_TMB_H_vs_L*100
#                                              ){
#   infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type_acronym == COI)),]
#   #infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type == COI)),]
# 
#   if(response_measure=='ORR'){
#     infunc_df_tmb=infunc_df_tmb[!is.na(infunc_df_tmb$ORR),]
#     threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
# 
#     # Contigency matrix
#     cont_matrix=table(infunc_df_tmb$TMB >= threshold, infunc_df_tmb$ORR)
#     # Perform fisher test
#     fisher_test_results_raw=fisher.test(cont_matrix)
#     to_return=c(effect_size=unlist(fisher_test_results_raw$estimate),
#                 P=fisher_test_results_raw$p.value)
#   } else if(response_measure =='OS') {
#     infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$Overall_survival_months),]
#     threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
# 
#     # Perform coxph test
#     coxph_model <- coxph(Surv(time = Overall_survival_months,
#                               event = Overall_survival_status) ~ (TMB >= threshold) ,
#                          data = infunc_df_tmb)
#     to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
#                 P=summary(coxph_model)$coefficients[1,c(5)])
#   }  else if(response_measure=='PFS') {
#     infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$PFS_months),]
#     threshold=sort(infunc_df_tmb$TMB, decreasing = T)[round(nrow(infunc_df_tmb)*(th/100))]
# 
#     coxph_model <- coxph(Surv(time = PFS_months,
#                               event = PFS_status) ~ (TMB >= threshold) ,
#                          data = infunc_df_tmb)
#     to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
#                 P=summary(coxph_model)$coefficients[1,c(5)])
#   }
#   names(to_return)=c('effect_size', 'P')
#   to_return
# }
```

```{r}
# Given a set of cancer types & Immuen matrix
# Tmb_Resp_Association - Or TRA
create_corr_with_Immune_factors<- function(infunc_cancer_types=allcancer_types,
                                           infunc_immune_matrix=immune_factor,
                                           infun_Tmb_Resp_Association = ORR_difference,
                                           response_type='ORR',
                                           consider_cancer_type='all_cancerTypes', 
                                           infunc_title='All cancerTypes'){
  infun_Tmb_Resp_Association=infun_Tmb_Resp_Association[
    !is.na(infun_Tmb_Resp_Association[,1]) & is.finite(infun_Tmb_Resp_Association[,1]),]
  infunc_cancer_types_filtered=intersect(rownames(infun_Tmb_Resp_Association),
                                         infunc_cancer_types)
  ## filter to remove features with high number of zeroes
  infunc_immune_matrix_filtered= infunc_immune_matrix[,!colSums(infunc_immune_matrix!= 0,na.rm = T) <= 8]
  
  ## Correlation
  filtered_features_resp = infunc_immune_matrix_filtered[na.omit(match(infunc_cancer_types_filtered,
                                                rownames(infunc_immune_matrix_filtered))),]
  
  target_resp= infun_Tmb_Resp_Association[rownames(filtered_features_resp),'effect_size']
  
  # cor.test(features_resp[,2],target_resp)
  correlation_matrix= data.frame(t(apply(filtered_features_resp, 2, function(x)
    unlist(cor.test(x, target_resp, method = 's')[c(3, 4)])) ))
  correlation_matrix=na.omit(correlation_matrix[order(correlation_matrix[,2],decreasing = T),])

  correlation_matrix$immune_factors=gsub('\\.',' ',gsub('\\.\\.',' ',rownames(correlation_matrix)))
    
  Panel2 <- ggplot(data=correlation_matrix, aes(x=reorder(immune_factors, estimate.rho),
                                                  y=estimate.rho
                                                  # , color= p.value <0.05
                                                  ))+
      geom_bar(aes(fill= estimate.rho),stat="identity")+
      scale_fill_gradient2(low = "red",mid = "dark grey",high = "dark green")+
      coord_flip() + # horizontal bars  
      geom_text(aes(y = 0, label = immune_factors, hjust = as.numeric(estimate.rho > 0))) +  # label text based on value
      theme_bw(base_size = 18)+
      theme(axis.text.y = element_blank(), legend.position="top",
            plot.title = element_text(color="red", size=14, face="bold.italic"))+
      labs(x='Immune Factors', y=paste('Spearman correlation with Power of TMB \n(using ',response_type,')',sep='' ))+
      ggtitle(infunc_title)
  ggsave(paste('../Result_fgures/tmb-responseVSTME_factors',
               consider_cancer_type,'_',
               response_type,'.pdf', sep ='' ),
         Panel2, height=7, width = 6, units = 'in', dpi = 500)
  list(correlation_matrix, Panel2)
}
```

```{r}
create_scatter_plot<- function(infunc_cancer_types=allcancer_types,
                               infunc_immune_matrix=immune_factor,
                               infun_Tmb_Resp_Association = OS_difference,
                               response_type='OS',
                               consider_cancer_type='all_cancerTypes', 
                               infunc_title='All cancerTypes',
                               immune_feature_of_interest= colnames(infunc_immune_matrix)[37]){
  df2plot=data.frame(infun_Tmb_Resp_Association[infunc_cancer_types,],
             features_resp= infunc_immune_matrix[infunc_cancer_types,immune_feature_of_interest])
  df2plot=na.omit(df2plot)
  Panel3 <- ggplot(df2plot, aes(x=rank(effect_size), y=rank(features_resp), size= -log(P, 10) )) + 
    stat_smooth(method='lm')+
    geom_point()+
    theme_bw(base_size = 15)+
    # theme(legend.position = 'none')+
    labs(x='Power of TMB (rank)',
         y=paste(gsub('\\.',' ',immune_feature_of_interest),  '(rank)'), size='-log10(P)')+
    stat_cor(label.y = 0, label.x = 5, method = 'p', cor.coef.name = 'rho', size=5)+
    geom_label_repel(aes(label=rownames(df2plot), size=0.5), show.legend = F)+
    coord_flip()
  ggsave(paste('../Result_fgures/tmb-responseVS',
               immune_feature_of_interest,
               consider_cancer_type,'_',
               response_type,'.pdf', sep ='' ),
         Panel3, height=5, width = 5, units = 'in', dpi = 500)
  # list(correlation_matrix, Panel3)
}
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


