---
title: "R Notebook"
## title: "Data preprocessing and function"
output: html_notebook
---
<!-- Libraries -->
```{r}
require(ggrepel)
library(forestplot)
library(corrplot)
require("survival")
require(statar)
library(ggplot2)
library(ggpubr)
library(readxl)
library(tidyr)
require("survival")
require("survminer")
library(caret)
#source('/Users/neelamf2/Downloads/Eytan/myCustom_functions.R')
# setwd('../data/')
```

<!--Load MSKCC combined data -->
```{r}
mskcc_combined= read.csv("../data/mskcc_combined.csv")
mskcc_combined$Cancer_Type_acronym=factor(mskcc_combined$Cancer_Type_acronym)

samples_by_cancer_type = aggregate(ID ~ Cancer_Type_acronym,mskcc_combined, length) 
cancerTypes_of_Interest = samples_by_cancer_type[,1][samples_by_cancer_type[,2]> 1]

## removing cancer types where PFS/ ORR is NA for figure 1B and ex figure 1
mskcc_df_ORR_PFS= mskcc_combined[!(is.na(mskcc_combined$PFS_status)),]
samples_each_cancer_type = aggregate(ID ~ Cancer_Type_acronym,mskcc_df_ORR_PFS, length) 
cancerTypes_of_Interest = samples_by_cancer_type[,1][samples_by_cancer_type[,2]> 1]

```

<!-- Functions needed-->
```{r}
# Calculates TMB and Response assocaition
Calculate_TMB_Response_association<-function(infunc_df= mskcc_combined,
                                             COI="NSCLC",
                                             response_measure='ORR'){
  infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type_acronym == COI)),]
  #infunc_df_tmb= infunc_df[((infunc_df$Cancer_Type == COI)),]
  
  if(response_measure=='ORR'){
    infunc_df_tmb=infunc_df_tmb[!is.na(infunc_df_tmb$ORR),]
    # Contigency matrix
    cont_matrix=table( infunc_df_tmb$TMB >=10, infunc_df_tmb$ORR)
    # Perform fisher test
    fisher_test_results_raw=fisher.test(cont_matrix)
    to_return=c(effect_size=unlist(fisher_test_results_raw$estimate),
                P=fisher_test_results_raw$p.value)
  } else if(response_measure =='OS') {
    infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$Overall_survival_months),]
    # Perform coxph test
    coxph_model <- coxph(Surv(time = Overall_survival_months,
                              event = Overall_survival_status) ~ (TMB >= 10) ,
                         data = infunc_df_tmb)
    to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
                P=summary(coxph_model)$coefficients[1,c(5)])
  }  else if(response_measure=='PFS') {
    infunc_df_tmb= infunc_df_tmb[!is.na(infunc_df_tmb$PFS_months),]
    coxph_model <- coxph(Surv(time = PFS_months,
                              event = PFS_status) ~ (TMB>=10) ,
                         data = infunc_df_tmb)
    to_return=c(effect_size=summary(coxph_model)$coefficients[1,c(2)],
                P=summary(coxph_model)$coefficients[1,c(5)])
  }
  names(to_return)=c('effect_size', 'P')
  to_return
}
```

```{r}
# Given a set of cancer types & Immuen matrix
# Tmb_Resp_Association - Or TRA
create_corr_with_Immune_factors<- function(infunc_cancer_types=allcancer_types,
                                           infunc_immune_matrix=immune_factor,
                                           infun_Tmb_Resp_Association = ORR_difference,
                                           response_type='ORR',
                                           consider_cancer_type='all_cancerTypes', 
                                           infunc_title='All cancerTypes'){
  infun_Tmb_Resp_Association=infun_Tmb_Resp_Association[
    !is.na(infun_Tmb_Resp_Association[,1]) & is.finite(infun_Tmb_Resp_Association[,1]),]
  infunc_cancer_types_filtered=intersect(rownames(infun_Tmb_Resp_Association),
                                         infunc_cancer_types)
  ## filter to remove features with high number of zeroes
  infunc_immune_matrix_filtered= infunc_immune_matrix[,!colSums(infunc_immune_matrix!= 0) <= 8]
  
  ## Correlation
  filtered_features_resp = infunc_immune_matrix_filtered[na.omit(match(infunc_cancer_types_filtered,
                                                rownames(infunc_immune_matrix_filtered))),]
  
  target_resp= infun_Tmb_Resp_Association[rownames(filtered_features_resp),'effect_size']
  
  # cor.test(features_resp[,2],target_resp)
  correlation_matrix= data.frame(t(apply(filtered_features_resp, 2, function(x)
    unlist(cor.test(x, target_resp, method = 's')[c(3, 4)])) ))
  correlation_matrix=na.omit(correlation_matrix[order(correlation_matrix[,2],decreasing = T),])

  correlation_matrix$immune_factors=gsub('\\.',' ',gsub('\\.\\.',' ',rownames(correlation_matrix)))
    
  Panel2 <- ggplot(data=correlation_matrix, aes(x=reorder(immune_factors, estimate.rho),
                                                  y=estimate.rho
                                                  # , color= p.value <0.05
                                                  ))+
      geom_bar(aes(fill= estimate.rho),stat="identity")+
      scale_fill_gradient2(low = "red",mid = "dark grey",high = "dark green")+
      coord_flip() + # horizontal bars  
      geom_text(aes(y = 0, label = immune_factors, hjust = as.numeric(estimate.rho > 0))) +  # label text based on value
      theme_bw(base_size = 18)+
      theme(axis.text.y = element_blank(), legend.position="top",
            plot.title = element_text(color="red", size=14, face="bold.italic"))+
      labs(x='Immune Factors', y=paste('Spearman correlation with Power of TMB \n(using ',response_type,')',sep='' ))+
      ggtitle(infunc_title)
  ggsave(paste('../Result_fgures/tmb-responseVSTME_factors',
               consider_cancer_type,'_',
               response_type,'.pdf', sep ='' ),
         Panel2, height=7, width = 6, units = 'in', dpi = 500)
  list(correlation_matrix, Panel2)
}
```

```{r}
create_scatter_plot<- function(infunc_cancer_types=allcancer_types,
                               infunc_immune_matrix=immune_factor,
                               infun_Tmb_Resp_Association = OS_difference,
                               response_type='OS',
                               consider_cancer_type='all_cancerTypes', 
                               infunc_title='All cancerTypes',
                               immune_feature_of_interest= colnames(infunc_immune_matrix)[37]){
  df2plot=data.frame(infun_Tmb_Resp_Association[infunc_cancer_types,],
             features_resp= infunc_immune_matrix[infunc_cancer_types,immune_feature_of_interest])
  df2plot=na.omit(df2plot)
  Panel3 <- ggplot(df2plot, aes(x=rank(effect_size), y=rank(features_resp), size= -log(P, 10) )) + 
    stat_smooth(method='lm')+
    geom_point()+
    theme_bw(base_size = 15)+
    # theme(legend.position = 'none')+
    labs(x='Power of TMB (rank)',
         y=paste(gsub('\\.',' ',immune_feature_of_interest),  '(rank)'), size='-log10(P)')+
    stat_cor(label.y = 0, label.x = 5, method = 'p', cor.coef.name = 'rho', size=5)+
    geom_label_repel(aes(label=rownames(df2plot), size=0.5), show.legend = F)+
    coord_flip()
  ggsave(paste('../Result_fgures/tmb-responseVS',
               immune_feature_of_interest,
               consider_cancer_type,'_',
               response_type,'.pdf', sep ='' ),
         Panel3, height=5, width = 5, units = 'in', dpi = 500)
  # list(correlation_matrix, Panel3)
}
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

